Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-07-04 Report/joint_wsrm_formulation.tex
16.48 Jul 4 2014

The \ac{WSRM} technique is predominantly used to design the precoders for the \ac{MIMO} \ac{BC} channels to maximize the weighted sum throughput in the downlink direction. The weights are used to provide different priorities to the users based on the respective \ac{QoS} criterion. Since the objective of this work is concerned with the reduction of the number of queued packets, we use the total number of the queued packets for each user as the respective weights. Using the size of the queued packets as the corresponding weights is obtained from the well known algorithm, called the \textit{backpreassure algorithm}, which is used to decide the routing decisions for each node in the network to efficiently send the packets to the desired destination with minimal packet drops and delay. The backpreassure algorithm is the outcome of minimizing the Lyapunov drift \cite{neely2010stochastic} conditioned over the current queue states \me{\mbf{Q}(i - 1)}, where \me{\mbf{Q}} is the vector formed by stacking the number of queued packets of each user in the system. Assuming the queue length grows according to \eqref{eqn-2a} for all the users, the Lyapunov drift conditioned over the current queue state is given by
\begin{equation} \label{eqn-3.1.0}
\Delta(i) = \frac{1}{2} \, \sum_{k \in \mc{U}} \, \mathrm{E} \Big{[}  \left ( Q_k(i)^2 - Q_k(i-1)^2 \right ) \vert \, \mbf{Q}(i - 1) \Big{]},
\end{equation}
where the expectation is taken over all possible control\footnote{transmission rates} decisions made in accordance to the channel under consideration to minimize the total backlogged packets at a given instant. Now, by substituting \eqref{eqn-2a} in \eqref{eqn-3.1.0}, the resulting expression is given by
\begin{equation}
\Delta(i) \leq \mathrm{E} \Big{[} \mbf{Q}(i-1) \left \lbrace \mbfa{\lambda}(i) - \mbf{t}(i) \right \rbrace | \mbf{Q}(i-1) \Big{]} + \frac{1}{2} \,\mathrm{E} \left [ \mbfa{\lambda}^2(i) | \mbf{Q}(i-1) \right ] + \frac{1}{2} \,\mathrm{E} \left [ \mbf{t}^2(i) | \mbf{Q}(i-1) \right ],
\label{drift-exp}
\end{equation}
where \me{\mbfa{\lambda}} and \me{\mbf{t}} are the vectors formed by stacking the arrival and transmission of all users in the system and the inequality provides the upper bound\footnote{\me{\left ( \max{\Big [ \mbf{Q}(i-1) - \mbf{t}(i),0 \Big ] } + \mbfa{\lambda}(i) \right )^2 \leq \mbf{Q}^2(i-1) + \mbf{t}^2(i) + \mbfa{\lambda}^2(i) + 2 \, \mbf{Q}(i-1) \left ( \mbfa{\lambda}(i) - \mbf{t}(i) \right ) }} for the Lyapunov drift.

The first term in \eqref{drift-exp} can be neglected from the optimization, since the arrivals are independent of the current queue state following \me{\mathrm{E} [ \mbf{Q}(i-1) \mbfa{\lambda}(i) | \mbf{Q}(i-1)] = \mbf{Q}(i-1)\mathrm{E} [\mbfa{\lambda}(i)] = \mbf{Q}(i-1)\mbf{A}(i)}. In the backpreassure algorithm, discussed in \cite{georgiadis2006resource,neely2010stochastic}, the second order terms are bounded by a constant \me{B}, \textit{i.e}, \me{\mathrm{E} \left [ \mbfa{\lambda}^2(i) | \mbf{Q}(i-1) \right ] + \mathrm{E} \left [ \mbf{t}^2(i) | \mbf{Q}(i-1) \right ] \leq B}. With this approximation, the resulting problem can be represented as
\begin{equation}
\underset{\mbf{t}(i)}{\text{minimize}} \sum_{k \in \mc{U}} \, Q_k(i-1) \: t_k(i).
\label{backpreassure}
\end{equation}
In \eqref{backpreassure}, the expectation operator is dropped since it is maximized by maximizing the function inside it. Now from \eqref{backpreassure}, the optimization variables \me{\mbf{t}(i)}, which is the vectorized transmission rates, are to be identified for each user for the current slot \me{i}. Since the current system can exploit both space and frequency resources through \ac{MIMO} \ac{OFDM} model, the transmission rates \me{\mbf{t}(i)} for the current instant can be controlled using the transmit precoders \me{\mbf{M}_{k,n}} designed for each user over each sub-channel resource. Since the decision variables \me{\mbf{t}(i)} depends on the queue state at \me{\mbf{Q}(i - 1)}, we drop the time variable \me{i} for the easy representation. The above problem defined in \eqref{backpreassure} can be rewritten to include the space-frequency allocation variable, namely \me{\mbf{M}_{k,n}}, as
\begin{IEEEeqnarray}{rCl} \label{q_gen_sum}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n} \\ t_{l,k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n} \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
\text{subject to.} & \quad & t_{l,k,n} \leq \log_2(1+\gamma_{l,k,n}) \IEEEyessubnumber \label{eqn-3.1.2} \\
& \quad &\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-3.1.3} \\
& \quad & \sum_{n=1}^N \sum_{l = 1}^L t_{l,k,n} \leq Q_k, \fall k \in \mc{U} \IEEEyessubnumber \label{eqn-3.1.4}
\end{IEEEeqnarray}
where \eqref{eqn-3.1.4} is due to the \me{[x]^+} operation in \eqref{eqn-2a} and \me{\gamma_{l,k,n}} is defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more number of backlogged packets, which essentially does the allocation in a greedy manner. As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as
\begin{IEEEeqnarray}{rCl}\label{gen_sum}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n} \\ t_{l,k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n} \IEEEyessubnumber \\
\text{subject to.} & \quad & \eqref{eqn-3.1.1}, \eqref{eqn-3.1.2}, \eqref{eqn-3.1.3} \; \text{and} \; \eqref{eqn-3.1.4}. \IEEEyessubnumber
\end{IEEEeqnarray}

>>>>

The queue minimizing algorithms are discussed extensively in the networking literature for providing congestion-free routing between any two nodes\footnote{refers to routers or PC's} in the network. It is achieved by finding a control policy considering the differential backlogs of the packets present in the source and the destination node, where control policy refers to the resource or rate allocation strategy. The control policy satisfying the condition is known as \emph{backpreassure algorithm}, discussed extensively in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm assumes the physical layer rate as a fixed entity due to the wired infrastructure. In order to use the algorithm in the wireless network, the user rate variable \me{t_{k}} should be modeled in accordance with the wireless backbone network.

The \ac{Q-WSRM} scheme is formulated by extending the \emph{backpreassure algorithm} for the \ac{MIMO} \ac{OFDM} network, where multiple \acp{BS} acts like a source nodes and the user terminals as a receiver nodes. The user rate variables \me{t_k} are obtained by the design of the transmit and the receive precoders to maximize the weighted sum rate. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order find the weights, following the approach in \cite{neely2010stochastic}, we bound the Lyapunov function\footnote{\me{\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q^2_k(i)}} drift as
\begin{IEEEeqnarray}{rCl}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot and the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
Now the conditional Lyapunov drift for the instant \me{i} is given by
\begin{IEEEeqnarray}{rCl}
\Delta(\mbf{Q}(i)) &\triangleq& \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the expectation depends on the control decisions made for the transmission variables \me{t_k} and the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}, the resulting problem can be given as
\begin{equation}
\underset{t_k}{\text{maximize}} \quad \sum_{k \in \mc{U}} \, Q_k(i) \, t_k(i),
\label{backpreassure}
\end{equation}
where the expectation operator is dropped since it is maximized by maximizing the function inside in an opportunistic manner.

Now the problem in \eqref{backpreassure} looks similar to the \ac{WSRM} problem if the weights are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, where the queue weighted sum rate maximization is considered as the objective function to determine the transmit precoders. In order to avoid the excessive  allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. With this, the \ac{Q-WSRM} problem for a wireless system is given by
\begin{IEEEeqnarray}{rCl} \label{q_gen_sum}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n} \\ t_{l,k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n} \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
\text{subject to.} & \quad & t_{l,k,n} \leq \log_2(1+\gamma_{l,k,n}) \IEEEyessubnumber \label{eqn-3.1.2} \\
& \quad &\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-3.1.3} \\
& \quad & \sum_{n=1}^N \sum_{l = 1}^L t_{l,k,n} \leq Q_k, \fall k \in \mc{U} \IEEEyessubnumber \label{eqn-3.1.4}
\end{IEEEeqnarray}
where the precoders are associated with the \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more number of backlogged packets, which essentially does the allocation in a greedy manner. As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as
\begin{IEEEeqnarray}{rCl}\label{gen_sum}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n} \\ t_{l,k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n} \IEEEyessubnumber \\
\text{subject to.} & \quad & \eqref{eqn-3.1.1}, \eqref{eqn-3.1.2}, \eqref{eqn-3.1.3} \; \text{and} \; \eqref{eqn-3.1.4}. \IEEEyessubnumber
\end{IEEEeqnarray}

### 2014-07-04 Report/kkt_mse_approach.tex
16.48 Jul 4 2014

In general, the distributed approaches presented in the earlier sections can be used to perform the precoder design problem in a decentralized way. One significant issue remains in both \acl{PD} and \ac{ADMM} apporaches is the update of the coupling variable. In both the methods, the coupling variables are updated by using the subgradients, which depends on the choice of the step size used for each update. In order to avoid the use of subgradient update for the coupling variables, the \ac{MSE} reformulation approach can be considered by solving the \ac{KKT} equations. Initial work on this approach is discussed extensively in \cite{wmmse_shi} for the \ac{WSRM} problem and \cite{kaleva2012weighted} considered the \ac{WSRM} problem with the additional \ac{QoS} constraints like minimum guaranteed rate requirements.

>>>>

In general, the distributed approaches presented in the earlier sections can be used to design the precoders in a decentralized way. The main issue affecting the convergence of the algorithm in both \acl{PD} and \ac{ADMM} apporaches is the update of the coupling variables using the step size. In order to avoid the use of subgradient update, the \ac{MSE} reformulation method is considered is studied extensively in the literature, such as, \cite{wmmse_shi} for the \ac{WSRM} problem and \cite{kaleva2012weighted} considered the \ac{WSRM} problem with the additional \ac{QoS} constraints like minimum guaranteed rate requirements.

### 2014-07-04 Report/problem_formulation_intro.tex
16.48 Jul 4 2014

In this paper, we address the problem of resource allocation across the \ac{MIMO} \ac{OFDM} framework, which are contended by the users served by multiple \acp{BS}. In order to minimize the number of queued packets, the precoders are designed to utilize the available space-frequency resources efficiently by limiting the over utilization of the same. In order to achieve this, we use the queue deviation utility as compared to the traditional \ac{WSRM} objective discussed extensively in the literature \cite{christensen2008weighted,wmmse_shi}. For practical and tractability reasons, we impose a constraint that the maximum number of transmitted bits for the user \me{k} is limited by the total backlogged packets available at the transmitter. As a result, the number of backlogged packets \me{v_k} remaining in the system for the user \me{k} is given by
\begin{equation}
v_k =  Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} \log_2(1+\gamma_{l,k,n}) \geq 0 \fall k \in \mathcal{U}.
\label{eqn-4.2}
\end{equation}
The above positivity constraint need to be satisified by \me{v_k} to avoid the excessive allocation of the resources.

The objective function for the precoder design problem is given by minimizing the absolute value of the queue deviation in \eqref{eqn-4.2} raised to the exponent \me{q} as \me{\text{minimize} \: \sum_{k \in \mc{U}} \, | v_k |^q}. With this objective, the problem of weighted queued packet minimization is given by the $q$-norm minimization as
\begin{IEEEeqnarray}{rCl}\label{eqn-3}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n}\\\mvec{W}{k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q\IEEEyessubnumber \\
\text{subject to} & \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-4.3}
\end{IEEEeqnarray}
where \me{\tilde{v}_k \triangleq a_k^{1/{q}}v_k}, \me{a_k} is the weighting factor which is incorporated to control user priority based on their respective \ac{QoS}, \me{\gamma_{l,k,n}} is defined in \eqref{eq:SINR}, \me{\mbf{M}_{k,n} \triangleq [ \, \mvec{m}{1,k,n} \, \mvec{m}{2,k,n} \dotsc \mvec{m}{L,k,n} \,]} comprises the beamformers associated with the user \me{k} for \me{n^\mathrm{th}} sub-channel transmission, and \me{\mbf{W}_{k,n} \triangleq [ \, \mvec{w}{1,k,n} \, \mvec{w}{2,k,n} \dotsc \mvec{w}{L,k,n} \,]} stacks the receive beamformers respectively. It can be easily extended for user specific streams \me{L_k} instead of using the common \me{L} streams for all users. In \eqref{eqn-4.3}, we consider the sum power constraint for each \ac{BS} across all sub-channels. Before discussing the solutions in detail, we discuss the existing algorithm to solve the issue of minimizing the number of backlogged packets with additional constraints required by problem.

>>>>

In this paper, we address the problem of resource allocation across the \ac{MIMO} \ac{OFDM} framework, in which the resources are contended by the \acp{BS} to empty the respective user queued packets. In order to minimize the number of queued packets, the precoders are designed to utilize the available space-frequency resources efficiently by limiting the over utilization of the same. To achieve this, we use the queue deviation utility as compared to the traditional \ac{WSRM} objective discussed extensively in the literature \cite{christensen2008weighted,wmmse_shi}. For practical and tractability reasons, we impose a constraint that the maximum number of transmitted bits for the user \me{k} is limited by the total backlogged packets available at the transmitter. As a result, the number of backlogged packets \me{v_k} remaining in the system for the user \me{k} is given by
\begin{equation}
v_k =  Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} \log_2(1+\gamma_{l,k,n}) \geq 0 \fall k \in \mathcal{U}.
\label{eqn-4.2}
\end{equation}
The above positivity constraint need to be satisified by \me{v_k} to avoid the excessive allocation of the resources.

The objective function for the precoder design problem is given by minimizing the absolute value of the queue deviation in \eqref{eqn-4.2} raised to the exponent \me{q} as \me{\text{minimize} \: \sum_{k \in \mc{U}} \, | v_k |^q}. With this objective, the problem of weighted queued packet minimization is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3}
\underset{\substack{\mvec{M}{k,n},\gamma_{l,k,n}\\\mvec{W}{k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q\IEEEyessubnumber \\
\text{subject to} & \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-4.3}
\end{IEEEeqnarray}
where \me{\tilde{v}_k \triangleq a_k^{1/{q}}v_k}, \me{a_k} is the weighting factor which is incorporated to control user priority based on their respective \ac{QoS}, \me{\gamma_{l,k,n}} is defined in \eqref{eq:SINR}, \me{\mbf{M}_{k,n} \triangleq [ \, \mvec{m}{1,k,n} \, \mvec{m}{2,k,n} \dotsc \mvec{m}{L,k,n} \,]} comprises the beamformers associated with the user \me{k} for \me{n^\mathrm{th}} sub-channel transmission, and \me{\mbf{W}_{k,n} \triangleq [ \, \mvec{w}{1,k,n} \, \mvec{w}{2,k,n} \dotsc \mvec{w}{L,k,n} \,]} stacks the receive beamformers respectively. It can be easily extended for user specific streams \me{L_k} instead of using the common \me{L} streams for all users. In \eqref{eqn-4.3}, we consider the sum power constraint for each \ac{BS} across all sub-channels. Before discussing the proposed solutions, we consider the existing algorithm to solve the issue of minimizing the number of backlogged packets with additional constraints required by problem.

### 2014-07-04 Report/system_model.tex
16.48 Jul 4 2014

In this paper we adopt linear beamforming technique at \acp{BS}. Specifically, the data symbols \me{d_{l,k,n}} for  user \me{k} on the \me{l^\mathrm{th}} spatial stream over the sub-channel \me{n} is multiplied with the beamformer \me{\mvec{m}{l,k,n} \in \mathbb{C}^{N_T \times 1}} for transmission.   In order to detect  multiple spatial streams at the receiver, a receive beamforming vector \me{\mvec{w}{l,k,n}} is employed at each user. Consequently, the received data symbol corresponding to the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n^\mathrm{th}} at user $k$ is given by
\begin{equation}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n}
\end{equation}
where \me{\mvec{H}{b,k,n} \in \mathbb{C}^{N_R \times N_T}} with rank \me{L = \min(N_R,N_T)} is the channel between the \ac{BS} \me{b} and  user \me{k} on the sub-channel \me{n}, and \me{\mvec{n}{k,n}} \me{\sim \mc{CN}(0,N_0)} is the additive noise vector for the user \me{k} on the \me{n^\mathrm{th}} sub-channel and \me{l^\mathrm{th}} spatial stream. Assuming  \me{||\mvec{w}{l,k,n}||_{2}^{2} = 1} and independent detection of data streams, we can write the signal-to-interference-plus-noise ratio (SINR) as
\begin{equation}\label{eq:SINR}
\gamma_{l,k,n} = \dfrac{\left |\mvec{w}{l,k,n}^\herm \, \mvec{H}{b_k,k,n} \, \mvec{m}{l,k,n} \right |^2}{N_0 + \sum_{(j,i) \neq (l,k)} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2}.
\end{equation}
With the infinite buffer model assumption, let \me{Q_k} be the number of backlogged packets which are destined for the user \me{k} at a given scheduling instant. The queue dynamics of the user \me{k} are modeled using the Poisson arrival process with the average packet arrivals of \me{A_k = \mathbf{E}_i\{\lambda_k\}} packets/bits, where \me{\lambda_k(i) \sim \mathrm{Pois}(A_k)} represents the instantaneous number of packets or bits arriving for the user \me{k} at the \me{{i}^\mathrm{th}} instant. The total number of queued packets at the \me{{(i+1)}^\mathrm{th}} instant for the user \me{k}, represented by \me{Q_k(i+1)}, follows
\begin{equation}
Q_k(i+1) = \Big [ Q_k(i) - t_k(i) \Big ]^+ + \lambda_k(i),
\label{eqn-2a}
\end{equation}
where \me{[x]^+ \equiv \max{ \lbrace x,0  \rbrace}} and \me{t_k} denotes the transmission in bits for user \me{k}. For a \ac{MIMO} \ac{OFDM} system,
\begin{equation}
t_k(i) = \sum_{n = 1}^N \, \sum_{l = 1}^L \, t_{l,k,n}(i),
\end{equation}
where \me{t_{l,k,n}} denotes the transmitted bits over \me{\ith{l}} spatial stream on the \me{\ith{n}} sub-channel. The maximum rate achieved over the \me{(l,n)} space-frequency resource is given by \me{\log_2(1 + \gamma_{l,k,n})}\footnote{assuming Gaussian signaling} for a given \ac{SINR} of \me{\gamma_{l,k,n}}. Note that the units of \me{t_k} and \me{Q_k} are in bits defined per channel use.

>>>>

In this paper we adopt linear beamforming technique at \acp{BS}. Specifically, the data symbols \me{d_{l,k,n}} for  user \me{k} on the \me{l^\mathrm{th}} spatial stream over the sub-channel \me{n} is multiplied with the beamformer \me{\mvec{m}{l,k,n} \in \mathbb{C}^{N_T \times 1}} for transmission.   In order to detect  multiple spatial streams at the receiver, a receive beamforming vector \me{\mvec{w}{l,k,n}} is employed at each user. Consequently, the received data symbol corresponding to the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n^\mathrm{th}} at user $k$ is given by
\begin{equation}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n}
\end{equation}
where \me{\mvec{H}{b,k,n} \in \mathbb{C}^{N_R \times N_T}} with rank \me{L = \min(N_R,N_T)} is the channel between the \ac{BS} \me{b} and  user \me{k} on the sub-channel \me{n}, and \me{\mvec{n}{k,n}} \me{\sim \mc{CN}(0,N_0)} is the additive noise vector for the user \me{k} on the \me{n^\mathrm{th}} sub-channel and \me{l^\mathrm{th}} spatial stream. Assuming  \me{||\mvec{w}{l,k,n}||_{2}^{2} = 1} and independent detection of data streams, we can write the signal-to-interference-plus-noise ratio (SINR) as
\begin{equation}\label{eq:SINR}
\gamma_{l,k,n} = \dfrac{\left |\mvec{w}{l,k,n}^\herm \, \mvec{H}{b_k,k,n} \, \mvec{m}{l,k,n} \right |^2}{N_0 + \sum_{(j,i) \neq (l,k)} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2}.
\end{equation}
With the infinite buffer model assumption, let \me{Q_k} be the number of backlogged packets which are destined for the user \me{k} at a given scheduling instant. The queue dynamics of the user \me{k} are modeled using the Poisson arrival process with the average packet arrivals of \me{A_k = \mathbf{E}_i\{\lambda_k\}} packets/bits, where \me{\lambda_k(i) \sim \mathrm{Pois}(A_k)} represents the instantaneous number of packets or bits arriving for the user \me{k} at the \me{{i}^\mathrm{th}} instant. The total number of queued packets at the \me{{(i+1)}^\mathrm{th}} instant for the user \me{k}, represented by \me{Q_k(i+1)}, follows
\begin{equation}
Q_k(i+1) = \Big [ Q_k(i) - t_k(i) \Big ]^+ + \lambda_k(i),
\label{eqn-2a}
\end{equation}
where \me{[x]^+ \equiv \max{ \lbrace x,0  \rbrace}} and \me{t_k} denotes the transmission in bits for user \me{k}. For a \ac{MIMO} \ac{OFDM} system,
\begin{equation}
t_k(i) = \sum_{n = 1}^N \, \sum_{l = 1}^L \, t_{l,k,n}(i),
\end{equation}
where \me{t_{l,k,n}} denotes the transmitted bits over \me{\ith{l}} spatial stream on the \me{\ith{n}} sub-channel. The maximum rate achieved over the \me{(l,n)} space-frequency resource is given by \me{\log_2(1 + \gamma_{l,k,n})}\footnote{assuming Gaussian signaling} for the \ac{SINR} of \me{\gamma_{l,k,n}}. Note that the units of \me{t_k} and \me{Q_k} are in bits defined per channel use.
